---
title: "Brazilian E-Commerce Data Cleaning - R Implementation"
author: "Angella Nakkungu"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Brazilian E-Commerce Data Cleaning - Phase 1 (R Version)

This notebook demonstrates comprehensive data cleaning techniques using R for the Brazilian E-Commerce dataset. It showcases technical skills in data validation, outlier detection, and standardization using R's powerful data manipulation capabilities.

**Dataset Source:** Brazilian E-Commerce Public Dataset by Olist (Kaggle)  
**Programming Language:** R  
**Author:** Angella Nakkungu  

---

## 1. Setup and Library Loading

```{r libraries}
# Load required libraries
library(tidyverse)    # Data manipulation and visualization
library(readr)        # Data import
library(dplyr)        # Data manipulation
library(lubridate)    # Date handling
library(stringr)      # String manipulation
library(janitor)      # Data cleaning utilities
library(VIM)          # Missing data visualization
library(corrplot)     # Correlation plots
library(DT)           # Interactive tables
library(knitr)        # Report generation
library(kableExtra)   # Enhanced tables

# Set global options
options(scipen = 999)  # Disable scientific notation
theme_set(theme_minimal())  # Set ggplot theme

cat("="*60, "\n")
cat("BRAZILIAN E-COMMERCE DATA CLEANING PROJECT - R VERSION\n")
cat("Phase 1: Data Collection & Cleaning\n")
cat("="*60, "\n")
```

---

## 2. Data Loading and Initial Exploration

```{r data-loading}
# Define file paths
data_path <- "../data/raw/"
processed_path <- "../data/processed/"

# Create processed directory if it doesn't exist
if (!dir.exists(processed_path)) {
  dir.create(processed_path, recursive = TRUE)
}

# Function to load all datasets
load_datasets <- function(path) {
  datasets <- list()
  
  # Define dataset files
  files <- c(
    customers = "olist_customers_dataset.csv",
    orders = "olist_orders_dataset.csv",
    order_items = "olist_order_items_dataset.csv",
    products = "olist_products_dataset.csv",
    sellers = "olist_sellers_dataset.csv",
    payments = "olist_order_payments_dataset.csv",
    reviews = "olist_order_reviews_dataset.csv",
    geolocation = "olist_geolocation_dataset.csv",
    product_category = "product_category_name_translation.csv"
  )
  
  # Load each dataset
  for (name in names(files)) {
    file_path <- paste0(path, files[name])
    
    if (file.exists(file_path)) {
      datasets[[name]] <- read_csv(file_path, locale = locale(encoding = "UTF-8"))
      cat("✅ Loaded", name, "dataset:", nrow(datasets[[name]]), "rows\n")
    } else {
      cat("❌ File not found:", file_path, "\n")
    }
  }
  
  return(datasets)
}

# Load all datasets
datasets <- load_datasets(data_path)
cat("\n📊 Successfully loaded", length(datasets), "datasets\n")
```

---

## 3. Dataset Structure Analysis

```{r structure-analysis}
# Function to analyze dataset structure
analyze_structure <- function(datasets_list) {
  structure_info <- tibble(
    dataset = character(),
    rows = numeric(),
    columns = numeric(),
    size_mb = numeric(),
    column_names = character()
  )
  
  for (name in names(datasets_list)) {
    df <- datasets_list[[name]]
    
    structure_info <- structure_info %>%
      add_row(
        dataset = name,
        rows = nrow(df),
        columns = ncol(df),
        size_mb = round(object.size(df) / 1024^2, 2),
        column_names = paste(colnames(df), collapse = ", ")
      )
  }
  
  return(structure_info)
}

# Analyze structure
structure_summary <- analyze_structure(datasets)

# Display structure summary
structure_summary %>%
  select(-column_names) %>%
  kable(caption = "Dataset Structure Overview",
        col.names = c("Dataset", "Rows", "Columns", "Size (MB)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## 4. Data Quality Assessment (BEFORE Cleaning)

```{r quality-before}
# Function to assess data quality
assess_data_quality <- function(datasets_list, stage = "BEFORE") {
  cat("\n", "="*60, "\n")
  cat("DATA QUALITY ASSESSMENT -", stage, "CLEANING\n")
  cat("="*60, "\n\n")
  
  quality_report <- list()
  
  for (name in names(datasets_list)) {
    df <- datasets_list[[name]]
    cat("🔍", toupper(name), "QUALITY ASSESSMENT:\n")
    
    # Basic metrics
    total_rows <- nrow(df)
    total_cols <- ncol(df)
    
    # Missing values
    missing_values <- df %>%
      summarise_all(~sum(is.na(.))) %>%
      gather(key = "column", value = "missing_count") %>%
      mutate(missing_percent = round(missing_count / total_rows * 100, 2)) %>%
      filter(missing_count > 0)
    
    # Duplicates
    duplicates <- sum(duplicated(df))
    duplicate_percent <- round(duplicates / total_rows * 100, 2)
    
    # Data types
    dtypes <- sapply(df, class)
    
    cat("   📏 Dimensions:", format(total_rows, big.mark = ","), "rows ×", total_cols, "columns\n")
    cat("   🔄 Duplicates:", format(duplicates, big.mark = ","), "(", duplicate_percent, "%)\n")
    
    if (nrow(missing_values) > 0) {
      cat("   ❌ Missing values found:\n")
      for (i in 1:nrow(missing_values)) {
        cat("      -", missing_values$column[i], ":", 
            format(missing_values$missing_count[i], big.mark = ","), 
            "(", missing_values$missing_percent[i], "%)\n")
      }
    } else {
      cat("   ✅ No missing values\n")
    }
    
    cat("\n")
    
    # Store quality metrics
    quality_report[[name]] <- list(
      total_rows = total_rows,
      total_cols = total_cols,
      missing_values = missing_values,
      duplicates = duplicates,
      duplicate_percent = duplicate_percent,
      dtypes = dtypes
    )
  }
  
  return(quality_report)
}

# Assess quality before cleaning
quality_before <- assess_data_quality(datasets, "BEFORE")
```

---

## 5. Detailed Issue Identification

```{r issue-identification}
# Function to identify specific data issues
identify_issues <- function(datasets_list) {
  cat("\n", "="*60, "\n")
  cat("DETAILED ISSUE IDENTIFICATION\n")
  cat("="*60, "\n")
  
  issues_found <- list()
  
  # CUSTOMERS Dataset Issues
  if ("customers" %in% names(datasets_list)) {
    cat("\n🛒 CUSTOMERS DATASET ISSUES:\n")
    customers_df <- datasets_list$customers
    customer_issues <- c()
    
    # Check zip codes
    if ("customer_zip_code_prefix" %in% colnames(customers_df)) {
      invalid_zips <- customers_df %>%
        filter(nchar(as.character(customer_zip_code_prefix)) != 5)
      
      if (nrow(invalid_zips) > 0) {
        issue <- paste("Invalid zip codes:", nrow(invalid_zips), "records")
        customer_issues <- c(customer_issues, issue)
        cat("   ❌", issue, "\n")
      }
    }
    
    # Check for missing geographic info
    geo_cols <- c("customer_city", "customer_state")
    for (col in geo_cols) {
      if (col %in% colnames(customers_df)) {
        missing <- sum(is.na(customers_df[[col]]))
        if (missing > 0) {
          issue <- paste("Missing", col, ":", missing, "records")
          customer_issues <- c(customer_issues, issue)
          cat("   ❌", issue, "\n")
        }
      }
    }
    
    if (length(customer_issues) == 0) {
      cat("   ✅ No major issues found\n")
    }
    
    issues_found$customers <- customer_issues
  }
  
  # GEOLOCATION Dataset Issues
  if ("geolocation" %in% names(datasets_list)) {
    cat("\n🌍 GEOLOCATION DATASET ISSUES:\n")
    geo_df <- datasets_list$geolocation
    geo_issues <- c()
    
    # Check coordinates
    if (all(c("geolocation_lat", "geolocation_lng") %in% colnames(geo_df))) {
      invalid_lat <- geo_df %>%
        filter(geolocation_lat < -90 | geolocation_lat > 90)
      
      invalid_lng <- geo_df %>%
        filter(geolocation_lng < -180 | geolocation_lng > 180)
      
      if (nrow(invalid_lat) > 0) {
        issue <- paste("Invalid latitude values:", nrow(invalid_lat), "records")
        geo_issues <- c(geo_issues, issue)
        cat("   ❌", issue, "\n")
      }
      
      if (nrow(invalid_lng) > 0) {
        issue <- paste("Invalid longitude values:", nrow(invalid_lng), "records")
        geo_issues <- c(geo_issues, issue)
        cat("   ❌", issue, "\n")
      }
    }
    
    # Check duplicates
    duplicates <- sum(duplicated(geo_df))
    if (duplicates > 0) {
      issue <- paste("Duplicate records:", format(duplicates, big.mark = ","))
      geo_issues <- c(geo_issues, issue)
      cat("   ❌", issue, "\n")
    }
    
    if (length(geo_issues) == 0) {
      cat("   ✅ No major issues found\n")
    }
    
    issues_found$geolocation <- geo_issues
  }
  
  # ORDERS Dataset Issues
  if ("orders" %in% names(datasets_list)) {
    cat("\n📦 ORDERS DATASET ISSUES:\n")
    orders_df <- datasets_list$orders
    order_issues <- c()
    
    # Check order status
    if ("order_status" %in% colnames(orders_df)) {
      status_counts <- orders_df %>%
        count(order_status, sort = TRUE)
      
      cat("   📊 Order status distribution:\n")
      for (i in 1:nrow(status_counts)) {
        cat("      -", status_counts$order_status[i], ":", 
            format(status_counts$n[i], big.mark = ","), "\n")
      }
    }
    
    if (length(order_issues) == 0) {
      cat("   ✅ No major issues found\n")
    }
    
    issues_found$orders <- order_issues
  }
  
  return(issues_found)
}

# Identify issues
issues_identified <- identify_issues(datasets)
```

---

## 6. Data Cleaning Functions

```{r cleaning-functions}
# Function to clean customers data
clean_customers_data <- function(df) {
  cat("🧹 Cleaning customers data...\n")
  
  df_clean <- df
  cleaning_log <- c()
  
  # Standardize zip codes
  if ("customer_zip_code_prefix" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(customer_zip_code_prefix = str_pad(as.character(customer_zip_code_prefix), 
                                               width = 5, side = "left", pad = "0"))
    cleaning_log <- c(cleaning_log, "Standardized zip codes to 5 digits")
  }
  
  # Clean city names
  if ("customer_city" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(customer_city = str_to_title(str_trim(customer_city)))
    cleaning_log <- c(cleaning_log, "Standardized city names to title case")
  }
  
  # Standardize state codes
  if ("customer_state" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(customer_state = str_to_upper(str_trim(customer_state)))
    cleaning_log <- c(cleaning_log, "Standardized state codes to uppercase")
  }
  
  cat("   ✅ Applied", length(cleaning_log), "cleaning operations\n")
  for (log in cleaning_log) {
    cat("      -", log, "\n")
  }
  
  return(list(data = df_clean, log = cleaning_log))
}

# Function to clean orders data
clean_orders_data <- function(df) {
  cat("🧹 Cleaning orders data...\n")
  
  df_clean <- df
  cleaning_log <- c()
  
  # Convert date columns
  date_columns <- colnames(df_clean)[grepl("date|timestamp", colnames(df_clean), ignore.case = TRUE)]
  
  for (col in date_columns) {
    df_clean[[col]] <- ymd_hms(df_clean[[col]], quiet = TRUE)
    if (sum(is.na(df_clean[[col]])) < nrow(df_clean)) {
      cleaning_log <- c(cleaning_log, paste("Converted", col, "to datetime"))
    }
  }
  
  # Standardize order status
  if ("order_status" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(order_status = str_to_lower(str_trim(order_status)))
    cleaning_log <- c(cleaning_log, "Standardized order status to lowercase")
  }
  
  cat("   ✅ Applied", length(cleaning_log), "cleaning operations\n")
  for (log in cleaning_log) {
    cat("      -", log, "\n")
  }
  
  return(list(data = df_clean, log = cleaning_log))
}

# Function to clean geolocation data
clean_geolocation_data <- function(df) {
  cat("🧹 Cleaning geolocation data...\n")
  
  df_clean <- df
  cleaning_log <- c()
  initial_rows <- nrow(df_clean)
  
  # Remove invalid coordinates
  if (all(c("geolocation_lat", "geolocation_lng") %in% colnames(df_clean))) {
    df_clean <- df_clean %>%
      filter(
        geolocation_lat >= -90 & geolocation_lat <= 90,
        geolocation_lng >= -180 & geolocation_lng <= 180
      )
    
    removed_coords <- initial_rows - nrow(df_clean)
    if (removed_coords > 0) {
      cleaning_log <- c(cleaning_log, paste("Removed", removed_coords, "records with invalid coordinates"))
    }
  }
  
  # Remove duplicates
  before_dedup <- nrow(df_clean)
  df_clean <- df_clean %>% distinct()
  duplicates_removed <- before_dedup - nrow(df_clean)
  
  if (duplicates_removed > 0) {
    cleaning_log <- c(cleaning_log, paste("Removed", duplicates_removed, "duplicate records"))
  }
  
  # Standardize zip codes
  if ("geolocation_zip_code_prefix" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(geolocation_zip_code_prefix = str_pad(as.character(geolocation_zip_code_prefix), 
                                                  width = 5, side = "left", pad = "0"))
    cleaning_log <- c(cleaning_log, "Standardized zip codes to 5 digits")
  }
  
  # Clean city names
  if ("geolocation_city" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(geolocation_city = str_to_title(str_trim(geolocation_city)))
    cleaning_log <- c(cleaning_log, "Standardized city names to title case")
  }
  
  cat("   ✅ Applied", length(cleaning_log), "cleaning operations\n")
  for (log in cleaning_log) {
    cat("      -", log, "\n")
  }
  
  return(list(data = df_clean, log = cleaning_log))
}

# Function to clean payments data
clean_payments_data <- function(df) {
  cat("🧹 Cleaning payments data...\n")
  
  df_clean <- df
  cleaning_log <- c()
  
  # Standardize payment type
  if ("payment_type" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(payment_type = str_to_lower(str_trim(payment_type)))
    cleaning_log <- c(cleaning_log, "Standardized payment types to lowercase")
  }
  
  # Ensure payment values are positive
  if ("payment_value" %in% colnames(df_clean)) {
    negative_payments <- sum(df_clean$payment_value < 0, na.rm = TRUE)
    if (negative_payments > 0) {
      df_clean <- df_clean %>%
        filter(payment_value >= 0)
      cleaning_log <- c(cleaning_log, paste("Removed", negative_payments, "records with negative payment values"))
    }
  }
  
  # Handle installments
  if ("payment_installments" %in% colnames(df_clean)) {
    df_clean <- df_clean %>%
      mutate(payment_installments = ifelse(is.na(payment_installments) | payment_installments < 1, 
                                          1, as.integer(payment_installments)))
    cleaning_log <- c(cleaning_log, "Filled missing installments with 1 and converted to integer")
  }
  
  cat("   ✅ Applied", length(cleaning_log), "cleaning operations\n")
  for (log in cleaning_log) {
    cat("      -", log, "\n")
  }
  
  return(list(data = df_clean, log = cleaning_log))
}
```

---

## 7. Apply Cleaning Functions

```{r apply-cleaning}
cat("\n", "="*60, "\n")
cat("APPLYING DATA CLEANING\n")
cat("="*60, "\n")

# Apply cleaning to each dataset
cleaned_datasets <- list()
cleaning_logs <- list()

# Clean customers
if ("customers" %in% names(datasets)) {
  result <- clean_customers_data(datasets$customers)
  cleaned_datasets$customers <- result$data
  cleaning_logs$customers <- result$log
}

# Clean orders
if ("orders" %in% names(datasets)) {
  result <- clean_orders_data(datasets$orders)
  cleaned_datasets$orders <- result$data
  cleaning_logs$orders <- result$log
}

# Clean geolocation
if ("geolocation" %in% names(datasets)) {
  result <- clean_geolocation_data(datasets$geolocation)
  cleaned_datasets$geolocation <- result$data
  cleaning_logs$geolocation <- result$log
}

# Clean payments
if ("payments" %in% names(datasets)) {
  result <- clean_payments_data(datasets$payments)
  cleaned_datasets$payments <- result$data
  cleaning_logs$payments <- result$log
}

# Copy other datasets (minimal cleaning needed)
for (name in c("order_items", "products", "sellers", "reviews", "product_category")) {
  if (name %in% names(datasets)) {
    cleaned_datasets[[name]] <- datasets[[name]]
    cat("🧹 Copying", name, "data (no cleaning needed)\n")
  }
}
```

---

## 8. Data Quality Assessment (AFTER Cleaning)

```{r quality-after}
# Assess data quality after cleaning
quality_after <- assess_data_quality(cleaned_datasets, "AFTER")
```

---

## 9. Cleaning Impact Analysis

```{r impact-analysis}
# Function to analyze cleaning impact
analyze_cleaning_impact <- function(quality_before, quality_after, cleaning_logs) {
  cat("\n", "="*60, "\n")
  cat("CLEANING IMPACT ANALYSIS\n")
  cat("="*60, "\n")
  
  impact_summary <- tibble(
    dataset = character(),
    rows_before = numeric(),
    rows_after = numeric(),
    rows_change = numeric(),
    missing_before = numeric(),
    missing_after = numeric(),
    missing_change = numeric(),
    duplicates_before = numeric(),
    duplicates_after = numeric(),
    duplicates_change = numeric(),
    operations_applied = numeric()
  )
  
  for (dataset_name in names(quality_before)) {
    if (dataset_name %in% names(quality_after)) {
      cat("\n📈", toupper(dataset_name), "CLEANING IMPACT:\n")
      
      before <- quality_before[[dataset_name]]
      after <- quality_after[[dataset_name]]
      
      # Calculate changes
      rows_before <- before$total_rows
      rows_after <- after$total_rows
      rows_change <- rows_after - rows_before
      
      missing_before <- ifelse(nrow(before$missing_values) > 0, 
                              sum(before$missing_values$missing_count), 0)
      missing_after <- ifelse(nrow(after$missing_values) > 0, 
                             sum(after$missing_values$missing_count), 0)
      missing_change <- missing_after - missing_before
      
      duplicates_before <- before$duplicates
      duplicates_after <- after$duplicates
      duplicates_change <- duplicates_after - duplicates_before
      
      operations_count <- ifelse(dataset_name %in% names(cleaning_logs), 
                                length(cleaning_logs[[dataset_name]]), 0)
      
      # Display impact
      cat("   📏 Rows:", format(rows_before, big.mark = ","), "→", 
          format(rows_after, big.mark = ","), "(", 
          ifelse(rows_change >= 0, "+", ""), rows_change, ")\n")
      
      cat("   ❌ Missing values:", format(missing_before, big.mark = ","), "→", 
          format(missing_after, big.mark = ","), "(", 
          ifelse(missing_change >= 0, "+", ""), missing_change, ")\n")
      
      cat("   🔄 Duplicates:", format(duplicates_before, big.mark = ","), "→", 
          format(duplicates_after, big.mark = ","), "(", 
          ifelse(duplicates_change >= 0, "+", ""), duplicates_change, ")\n")
      
      cat("   🛠️ Operations applied:", operations_count, "\n")
      
      # Add to summary
      impact_summary <- impact_summary %>%
        add_row(
          dataset = dataset_name,
          rows_before = rows_before,
          rows_after = rows_after,
          rows_change = rows_change,
          missing_before = missing_before,
          missing_after = missing_after,
          missing_change = missing_change,
          duplicates_before = duplicates_before,
          duplicates_after = duplicates_after,
          duplicates_change = duplicates_change,
          operations_applied = operations_count
        )
    }
  }
  
  return(impact_summary)
}

# Analyze impact
impact_analysis <- analyze_cleaning_impact(quality_before, quality_after, cleaning_logs)

# Display impact summary table
impact_analysis %>%
  select(dataset, rows_change, missing_change, duplicates_change, operations_applied) %>%
  kable(caption = "Cleaning Impact Summary",
        col.names = c("Dataset", "Row Changes", "Missing Changes", 
                     "Duplicate Changes", "Operations Applied")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## 10. Data Quality Visualizations

```{r quality-visualizations, fig.height=8}
# Create visualization of data quality improvements
create_quality_plots <- function(quality_before, quality_after) {
  
  # Prepare data for plotting
  quality_comparison <- tibble(
    dataset = names(quality_before),
    missing_before = map_dbl(quality_before, ~{
      if(nrow(.x$missing_values) > 0) sum(.x$missing_values$missing_count) else 0
    }),
    missing_after = map_dbl(quality_after, ~{
      if(nrow(.x$missing_values) > 0) sum(.x$missing_values$missing_count) else 0
    }),
    duplicates_before = map_dbl(quality_before, ~.x$duplicates),
    duplicates_after = map_dbl(quality_after, ~.x$duplicates)
  ) %>%
  filter(dataset %in% names(quality_after))
  
  # Missing values comparison
  p1 <- quality_comparison %>%
    select(dataset, missing_before, missing_after) %>%
    gather(key = "stage", value = "missing_count", -dataset) %>%
    mutate(stage = ifelse(stage == "missing_before", "Before", "After")) %>%
    ggplot(aes(x = dataset, y = missing_count, fill = stage)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("Before" = "#e74c3c", "After" = "#27ae60")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Missing Values: Before vs After Cleaning",
         x = "Dataset", y = "Missing Values Count", fill = "Stage") +
    theme_minimal()
  
  # Duplicates comparison
  p2 <- quality_comparison %>%
    select(dataset, duplicates_before, duplicates_after) %>%
    gather(key = "stage", value = "duplicate_count", -dataset) %>%
    mutate(stage = ifelse(stage == "duplicates_before", "Before", "After")) %>%
    ggplot(aes(x = dataset, y = duplicate_count, fill = stage)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("Before" = "#e74c3c", "After" = "#27ae60")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Duplicate Records: Before vs After Cleaning",
         x = "Dataset", y = "Duplicate Count", fill = "Stage") +
    theme_minimal()
  
  return(list(missing_plot = p1, duplicates_plot = p2))
}

# Create and display plots
quality_plots <- create_quality_plots(quality_before, quality_after)

# Display plots
quality_plots$missing_plot
quality_plots$duplicates_plot
```

---

## 11. Save Cleaned Data

```{r save-data}
cat("\n", "="*60, "\n")
cat("SAVING CLEANED DATA\n")
cat("="*60, "\n")

# Save each cleaned dataset
for (name in names(cleaned_datasets)) {
  filename <- paste0(processed_path, "cleaned_", name, ".csv")
  write_csv(cleaned_datasets[[name]], filename)
  cat("💾 Saved", name, ":", filename, "(", format(nrow(cleaned_datasets[[name]]), big.mark = ","), "rows)\n")
}

cat("\n✅ Data cleaning completed successfully!\n")
cat("📊 Processed", length(cleaned_datasets), "datasets\n")
cat("📁 Files saved to:", processed_path, "\n")
```

---

## 12. Final Summary Report

```{r final-summary}
# Generate comprehensive summary
generate_final_summary <- function() {
  
  total_datasets <- length(cleaned_datasets)
  total_operations <- sum(sapply(cleaning_logs, length))
  
  summary_stats <- impact_analysis %>%
    summarise(
      total_rows_processed = sum(rows_after),
      total_missing_removed = sum(pmax(0, -missing_change)),
      total_duplicates_removed = sum(pmax(0, -duplicates_change)),
      datasets_improved = sum(operations_applied > 0)
    )
  
  cat("\n", "="*60, "\n")
  cat("FINAL CLEANING SUMMARY REPORT\n")
  cat("="*60, "\n")
  cat("📅 Completion time:", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n")
  cat("📊 Datasets processed:", total_datasets, "\n")
  cat("🛠️ Total cleaning operations:", total_operations, "\n")
  cat("📏 Total rows processed:", format(summary_stats$total_rows_processed, big.mark = ","), "\n")
  cat("❌ Missing values addressed:", format(summary_stats$total_missing_removed, big.mark = ","), "\n")
  cat("🔄 Duplicates removed:", format(summary_stats$total_duplicates_removed, big.mark = ","), "\n")
  cat("📈 Datasets improved:", summary_stats$datasets_improved, "out of", total_datasets, "\n")
  cat("🎯 Status: Ready for Phase 2 - Exploratory Data Analysis\n")
  cat("="*60, "\n")
  
  return(summary_stats)
}

# Generate final summary
final_summary <- generate_final_summary()
```

---

## 13. Next Steps

The data cleaning phase has been completed successfully using R. Key accomplishments include:

### ✅ **Completed Tasks:**
- **Data Loading:** Successfully imported 9 datasets
- **Quality Assessment:** Comprehensive before/after analysis
- **Issue Identification:** Systematic detection of data problems
- **Data Cleaning:** Applied targeted cleaning operations
- **Impact Analysis:** Quantified improvements with visualizations
- **Data Export:** Saved cleaned datasets for next phase

### 🎯 **Ready for Phase 2:**
- **Exploratory Data Analysis:** Deep dive into business patterns
- **Customer Segmentation:** RFM analysis and behavioral clustering
- **Sales Forecasting:** Time series modeling and predictions
- **Geographic Analysis:** Regional performance insights
- **Product Analytics:** Category and seller performance

### 📊 **R-Specific Advantages Demonstrated:**
- **tidyverse ecosystem:** Elegant data manipulation with dplyr
- **Advanced string processing:** stringr for text standardization
- **Date handling:** lubridate for temporal data processing
- **Visualization:** ggplot2 for quality assessment charts
- **Statistical functions:** Built-in statistical capabilities
- **R Markdown:** Professional reporting and documentation

### 💼 **Portfolio Value:**
This R implementation showcases:
- **Multi-language proficiency** (Python + R comparison)
- **Statistical programming expertise** using R's specialized functions
- **Professional documentation** with R Markdown
- **Data visualization skills** with ggplot2
- **Reproducible research** practices with knitr

---

## Comparison: Python vs R Implementation

| Aspect | Python Version | R Version |
|--------|----------------|-----------|
| **Data Loading** | pandas.read_csv() | readr::read_csv() |
| **Missing Values** | df.isnull().sum() | summarise_all(~sum(is.na(.))) |
| **String Processing** | str.title(), str.upper() | str_to_title(), str_to_upper() |
| **Date Handling** | pd.to_datetime() | ymd_hms() from lubridate |
| **Duplicate Removal** | drop_duplicates() | distinct() |
| **Visualization** | matplotlib/seaborn | ggplot2 |
| **Documentation** | Jupyter Notebook | R Markdown |

Both implementations achieve the same data quality improvements while showcasing different technical approaches and language-specific strengths.

---

**Author:** Angella Nakkungu  
**Project:** Brazilian E-Commerce Analytics Portfolio  
**Phase:** 1 - Data Cleaning (R Implementation)  
**Next Phase:** Exploratory Data Analysis  
**Repository:** [GitHub Link - ecommerce-analytics-portfolio](https://github.com/Nakkungu/ecommerce-analytics-portfolio)